Relationships in the extremes and their influence on competition and coexistence

### 2021-06-23 paper/Paper.Rnw
further progress on theory section

%attempt 3 at nat and sharp
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,.333ex) -- (.666ex,.333ex);%
\draw (.333ex,.666ex) -- (.666ex,.666ex);%
\draw (.333ex,.333ex) -- (.333ex,1ex);%
\draw (.66ex,0) -- (.66ex,.66ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

@@

For pedgogical clarity, we develop each step of the theory both for a simple classical model 
(the lottery model), and in general. The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
We denote $\cor(b_1, b_2)$ by $\rho$. However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

@@

Having just reviewed storage effects, we now specify how to decompose storage effects into
contrbutions due to asymmetric tail associations and contributions due
to $EC$ covariance \emph{per se}. The storage effect for species $i$ was computed by comparing
a difference between invader and resident mean growth rates for the original model to the same quantity calculated for a 
modified model for which $E_i$ and $C_i$ were rendered independent (see (\ref{eq:storage_effects_gen})). 
Because $E_i$ and $C_i$ were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole relationship between $E_i$ and $C_i$, in all its aspects, to the invader/resident growth rate
difference. However, we can instead consider a distribution $(E_i^\nat,C_i^\nat)$ with the same
marginal distributions as $(E_i,C_i)$ and with some but not all aspects of the relationship 
between the components modified relative to $(E_i,C_i)$; specifically, tail associations are rendered symmetric
(see below for how this is done). We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_i^\nat)$,
and we consider the decomposition 
\begin{align}
\Delta I_i &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \\
&= I_i^{(\text{ata})} + I_i^{(\text{cps})}.\label{eq:storage_effects_further}
\end{align}
The superscript ``ata" stands for ``asymmetric tail associations'' because this component of storage effects is due to 
asymmetric tail associations between $E_i$ and $C_i$. The superscript ``cps" stands for ``covariance
\emph{per se}'' because this component of storage effects is due to $EC$ covariance \emph{per se}.

>>>>

%attempt 3 at nat and sharp
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

@@

For pedgogical clarity, we develop each step of the theory both for a simple classical model 
(the lottery model), and in general. The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

@@

Having just reviewed storage effects, we now specify how to decompose storage effects into
contrbutions due to asymmetric tail associations and contributions due
to $EC$ correlation \emph{per se}. The storage effect for species $i$ was computed by comparing
a difference between invader and resident mean growth rates for the original model to the same quantity calculated for a 
modified model for which $E_i$ and $C_i$ were rendered independent (see (\ref{eq:storage_effects_gen})). 
Because $E_i$ and $C_i$ were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole relationship between $E_i$ and $C_i$, in all its aspects, to the invader/resident growth rate
difference. However, we can instead consider a distribution $(E_i^\nat,C_i^\nat)$ with the same
marginal distributions as $(E_i,C_i)$ and with some but not all aspects of the relationship 
between the components modified relative to $(E_i,C_i)$; specifically, asymmetries of tail association are removed
(see below for how this is done). We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_i^\nat)$,
and we consider the decomposition 
\begin{align}
\Delta I_i &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \\
&= I_i^{(\text{ata})} + I_i^{(\text{cps})}.\label{eq:storage_effects_further}
\end{align}
The superscript ``ata" stands for ``asymmetric tail associations'' because this component of storage effects is due to 
asymmetries of tail associations between $E_i$ and $C_i$. The superscript ``cps" stands for ``correlation
\emph{per se}'' because this component of storage effects is due to $EC$ correlation \emph{per se}.
The notation $\nat$ was chosen because the new symbol comprises part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the fact that part of the relationship between $E$ and $C$ components is removed.

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? We assume that the cummulative distribution functions (cdfs) of $E_i$ and 
$C_i$, which we denote $F_{E_i}$ and $F_{C_i}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the cdf of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_i}(C_i)$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal.  Letting $\rho_i$ be the covariance of
these random variables, we then define a bivariate normal random variable $(e_i,c_i)$ such that
$\var(e_i)=\var(c_i)=1$ and such that $\cov(e_i,c_i)=\tilde{\rho}_i$. We define $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_i^\nat=F_{C_i}^{-1} \circ \varphi(c_i)$. The value of $\tilde{\rho}_i$ is selected in one of 
four alternative ways: $\tilde{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_i^\nat$ matches that of $E_i$ and $C_i$. The choice of bivariate normal 
$(e_i,c_i)$ ensures that $E_i^\nat$ and $C_i^\nat$ are symmetrically tail associated in the sense
described in the Introduction.
%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
%the *right* way
For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
we have replaced the copula of $(E_i,C_i)$ with a normal copula to form $(E_i^\nat,C_i^\nat)$.
The choice of $\tilde{\rho}_i$ ensures that $E_i^\nat$ and $C_i^\nat$ are correlated to the same extent as 
$E_i$ and $C_i$, according to whichever way of defining correlation is preferable to the researcher. Thus $E_i^\nat$ and 
$C_i^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired.

%***DAN: next step - work it all out for the lottery model now
